---
title: "Toy example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Toy example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results = F}
library(bayesEpi)
library(ggplot2)
# devtools::load_all("~/Git/bayesEpi")
# devtools::install_github("https://github.com/Smoothing-IWP/OSplines/tree/OSpline-Original") # If wanna use o-splines...
```

This vignette shows hot to use the `fitModel` function to fit a case crossover model.

## Toy data
Consider the following dummy data:
```{r data}
set.seed(666)
n <- 1500
toyData <- data.frame(time = seq(n))

# explanatory variables
toyData$hum <- pmin(pmax((sin(toyData$time/365 * 2*pi) + 1)/2 + rnorm(n,0,.75),0),1)
toyData$o3 <- pmax((sin(toyData$time/365 * 2*pi) + 1)*25 + rnorm(n,0,10),0)

# overdispersion and hazard
theta <- 6
f_hum <- function(x) abs(1.25-x)
f_o3 <- function(x) sqrt(x/100)
f_time <- function(x) .5 * sin(2*pi * x/365.25) + log(1+x/n) 

# response variable
toyData$count <- rpois(n,exp(f_hum(toyData$hum) + f_o3(toyData$o3) + f_time(toyData$time) + rnorm(n, 0, exp(-theta/2))))

head(toyData)
```
Here, the names `hum` and `o3` refer to humidity level and ozone concentration, which are two common variables in air pollution epidemiology (although the values they take here do not necessarily make sense).

## Model specification
To fit a case crossover model with `count` as response variable, `time` as the time index, `hum` as fixed effect and `o3` as random effect, we first specify the model via
```{r model}

# fixed effect model via splines (somewhat tricky to set up the knots, two sets for each side of the reference values)
# model <- bayesEpi::ccModel(response = "count", time_index = "time",
#                         fixed = list(
#                            "hum" = fixedEffect(bayesEpi:::poly_effect(ref_value = median(toyData$hum))),
#                            "o3" = fixedEffect(bayesEpi:::bs_effect(knots = list(c(0,0,15,30,30), c(30,30,45,60,76,76)),
#                                                                    degree = 3, ref_value = 30))),
#                          random = NULL,
#                          design = ccDesign(scheme = "time stratified", n_control=3),
#                          overdispersion = randomEffect(gaussian_effect(), log_gamma_prior(shape=.5, rate=1e-7)),
#                          aghq_input = aghqInput())

# rw2 model
model <- bayesEpi::ccModel(response = "count", time_index = "time",
                           fixed = list("hum" = fixedEffect(bayesEpi:::poly_effect(ref_value = median(toyData$hum)))),
                           random = list("o3" = randomEffect(rw_effect(ref_value = 25, order=2,  binwidth=1), 
                                                             pc_prec_prior(alpha=.5, u=.05))),
                           design = ccDesign(scheme = "time stratified", n_control=3, lag=1),
                           overdispersion = randomEffect(gaussian_effect(), pc_prec_prior(alpha=.5, u=.03)),
                           aghq_input = aghqInput())

```
More fixed or random effects could be added, e.g., with
`random = list("o3" = randomEffect(rw_effect(), pc_prep_prior()), "other_random_effect" = randomEffect(rw_effect(), pc_prep_prior()))`.


## Fit
We can then fit the model with the function `fitModel`, which usually prints a bunch of stuff (somewhat uninformative if you know nothing about the `aghq` package).
```{r fit, results = F}
fit <- fitModel(model, toyData, silent = T)
```

## Analysis of the results
Some functions are provided for analyzing the results returned by `fitModel`.
Median prediction as well as credible intervals can be obtained via the function `getResults`.
```{r getResults, results = T}
# values at which to evaluate spline, if used
# values <- list("o3" = seq(0,75,1))
values <- NULL # will use bins for rw2

res <- getResults(fit, values = values, 
                  M = 5000, # M is number of samples to draw (for inference)
                  probs_pw = c(.9), # nominal levels for pointwise CI
                  probs_g = c(.9)) # nominal levels for global envelopes
res[res$parameter_type == "theta",]
```


You can plot them as follows
```{r plots, results = F}
# subset results (and construct true effect data frame, for comparison)
# if using splines model
# res0 <- res[res$parameter_type == "beta*" & res$variable_name == "o3",]
# ref_val <- fit$model$fixed$o3$model$params$ref_value

# if using rw2 model
res0 <- res[res$parameter_type == "gamma*" & res$variable_name == "o3",]
values <- list("o3" = fit$model$random$o3$model$extra$bin_values)
ref_val <- fit$model$random$o3$model$extra$rounded_ref_value

true_effect <- data.frame(
  variable_name = "o3",
  variable_value = values$o3,
  median = f_o3(values$o3) - f_o3(ref_val)
)


ggplot(res0, aes(x=variable_value, y=median)) +
  theme_bw() +
  xlab("exposure (o3)") + ylab("log-relative risk") +
  geom_line(data=true_effect, linetype=2) +
  geom_vline(xintercept = ref_val, linetype = 3) +
  geom_line() +
  geom_ribbon(aes(ymin=ge_lo_90, ymax=ge_hi_90), alpha = .25)

```

We can also take a look at the overdispersion coefficients:
```{r, overdispersion plot, results = F}

res0 <- getResults_z(fit, probs_pw = .9, probs_g = NULL, M = 5000)

ggplot(res0, aes(x=variable_value, y=median)) +
  theme_bw() +
  xlab("overdispersion coef. (z)") + ylab("log-relative risk") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(size=.35) +
  geom_ribbon(aes(ymin=perc_5, ymax=perc_95), alpha = .25)

```
