---
title: "Toy example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Toy example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results = F}
#library(bayesEpi)
devtools::load_all("~/Git/packages/bayesEpi")
```

This vignette shows hot to use the `fitModel` function to fit a case crossover model.

## Toy data
Consider the following dummy data:
```{r data}
n <- 1000
toyData <- data.frame(time = seq(n))

# explanatory variables and hazard
toyData$hum <- pmin(pmax((sin(toyData$time/365 * 2*pi) + 1)/2 + rnorm(n,0,.75),0),1)
toyData$o3 <- pmax((sin(toyData$time/365 * 2*pi) + 1)*25 + rnorm(n,0,10),0)
loghazard <- abs(1.25-toyData$hum)*1 + (toyData$o3/25)^2

# response variable
toyData$count <- rpois(n,exp(loghazard))

head(toyData)
```
Here, the names `hum` and `o3` refer to humidity level and ozone concentration, which are two common variables in air pollution epidemiology (although the values they take here do not necessarily make sense).

## Model specification
To fit a case crossover model with `count` as response variable, `time` as the time index, `hum` as fixed effect and `o3` as random effect, we first specify the model via
```{r model}
model <- bayesEpi::ccModel(response = "count",
                           time_index = "time",
                           fixed = list("hum" = fixedEffect(gaussian_prior())),
                           random = list("o3" = randomEffect(rw_effect(), pc_prep_prior())),
                           overdispersion = randomEffect(gaussian_effect(), pc_prep_prior()),
                           design = ccDesign(),
                           control_aghq = controlAGHQ())
```
More fixed or random effects could be added, e.g., with
`random = list("o3" = randomEffect(rw_effect(), pc_prep_prior()), "other_random_effect" = randomEffect(rw_effect(), pc_prep_prior()))`.
Note that only the gaussian prior is implemented for fixed effects, and that only the random walk model (`rw_effect`) is implemented for random effects.

## Fit
We can then fit the model with the function `fitModel`, which usually prints a bunch of stuff (somewhat uninformative if you know nothing about the `aghq` package).
```{r fit, results = F}
fit <- fitModel(model, toyData)
```

## Analysis of the results
Several functions are provided for analyzing the results returned by `fitModel`.
Median prediction as well as confidence intervals can be obtained via the function `getResults`.
```{r results, results = F}
# res <- getResults(fit, quantiles = c(.05,.5,.95))
```

```{r plots, results = F}
# plotResults(res, alpha = .1)
```
